---
title: " WDS HW 3"
author:
- Group Member 1
- Group Member 2
- Group Member 3
- Group Member 4
- Group Member 5
date: 'Due: 10:00PM,  July 21, 2021'
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "hide", fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output
# check if you have ISLR package, if not, install it
if(!require('pacman')) {install.packages('pacman')}
pacman::p_load(ISLR, readxl, tidyverse, magrittr, dplyr, ggplot2, leaps, car, contrast) 
# ISLR will be installed and the that AUTO needed for this hw will be there
```


\pagebreak



# Case study: Automobiles efficiency

Are cars being built more efficient? Are Asian cars more efficient than cars built in America or Europe? To answer the questions we will use the `Auto` dataset from ISLR. The original dataset contains 408 observations about cars. It is similar to the CARS dataset that we use in our lectures. But it also collects information by years. To get the data, first install the package ISLR. The `Auto` dataset should be loaded automatically. The original data source is here: https://archive.ics.uci.edu/ml/datasets/auto+mpg

Get familiar with this dataset first. A good data set should be well documented. Use the command `?ISLR::Auto` to view a description of the dataset. Please add the variable list with names, brief descriptions and units of the variables below. 

```{r}
data1 <- Auto
str(data1)
names(data1)
summary(data1)
```

# EDA

Explore the data first. 

i. What is the range of `year`? Why is this important to know?  

From the summary above, we can see that year ranges from 1970 to 1982. (actually 70-82 in the dataset). This lets us know the extent to which our data carries over into modern data.

ii. Should `origin` be a continuous variable? Why or why not. In any case make `origin` a categorical variable. 

It should be categorical as they denote places of origin (American, European, Japanese respectively). 

``` {r make_categorical}
# Make the origin col categorical
data1[data1$origin == 1, c("origin")] <- "American"
data1[data1$origin == 2, c("origin")] <- "European"
data1[data1$origin == 3, c("origin")] <- "Japanese"
data1$origin <- as.factor(data1$origin)
data1$cylinders <- as.factor(data1$cylinders)
```

iii. Do you see any peculiarity in the data?  

The number of American cars far outweighs the number of European or Japanese cars.


## What effect does `time` have on `MPG`?

i. Show a scatter plot of `mpg` vs. `year` with the LS line imposed. Does the plot show a positive trend? 

Yes the plot shows that mpg increases over time on average.

``` {r mpg_year_plot}

fit1 <- lm(mpg~year, data = data1)
ggplot(data1, aes(x = year, y = mpg)) +geom_point()+geom_smooth(method = "lm", formula = 'y~x', se=F)

```

ii. Now run a simple regression of `mpg` vs. `year` and report R's `summary` output. Is `year` a significant variable at the .05 level? State what effect `year` has on `mpg`, if any, according to this model. 

Year is significant at the .05 level. On average, the MPG increases by 1.23 per year.

``` {r mpg_year_fit}

fit1 <- lm(mpg ~ year, data = data1)
summary(fit1)

```

iii. Add `horsepower` on top of the variable `year` to your linear model. Is `year` still a significant variable at the .05 level? Give a precise interpretation of the `year`'s effect found here. 

Year is still significant. On average, cars with the same horsepower see an average increase of 0.657 MPG per year.

``` {r hp_mpg_yr_fit}

fit2 <- lm(mpg ~ year + horsepower, data = data1)
summary(fit2)

```

iv. The two 95% CI's for the coefficient of year differ among ii. and iii. How would you explain the difference to a non-statistician?

The two differ because holding horsepower constant changes the effect of year.

v. Create a model with interaction by fitting `lm(mpg ~ year * horsepower)`. Is the interaction effect significant at .05 level? Explain the year effect (if any). 

The interaction is significant at .05 level. Here the mpg for a given horsepower H increases by 2.19 - 0.016H per year.

``` {r}

fit3 = lm(mpg ~ year * horsepower, data = data1)
summary(fit3)
reg1 <- lm(mpg~year * horsepower,data=data1) 
	summary(reg1)
	with(data1,plot(year, mpg))
	abline(reg1)
```

<!-- ## Categorical predictors -->

<!-- Remember that the same variable can play different roles! Take a quick look at the variable `cylinders`, and try to use this variable in the following analyses wisely. We all agree that a larger number of cylinders will lower mpg. However, we can interpret `cylinders` as either a continuous (numeric) variable or a categorical variable. -->

<!-- a) Fit a model that treats `cylinders` as a continuous/numeric variable. Is `cylinders` significant at the 0.01 level? What effect does `cylinders` play in this model? -->

<!-- b) Fit a model that treats `cylinders` as a categorical/factor. Is `cylinders` significant at the .01 level? What is the effect of `cylinders` in this model? Describe the `cylinders` effect over `mpg`.  -->

<!-- c) What are the fundamental differences between treating `cylinders` as a continuous and categorical variable in your models?  -->


## Bring origin into the model

i. Do mpg's differ on average among different origin? Fit a linear model with mpg vs. origin. Report the output. 

MPGs do differ on average. From the model we see that on average European cars have 7.57 more mpg than American cars and Japanese cars have an average of 10.42 more MPG than American cars. Japanese cars have on average 2.85 more MPG than European cars. (For full summaries, see code for next section)

``` {r}

fit_origin = lm(mpg ~ origin, data = data1)
summary(fit_origin)

```

a) Are mpg's on average different among three regions? Perform a test at .01 level. When you reject the null hypothesis, what have you proved? 

There are notable differences between the different continents at the 0.1 level. The null hypothesis of there being no significant difference can be rejected from the results of the Anova call below. This shows that the 3 origins are all different.

```{r}

# data1$origin <- factor(data1$origin, levels = c("American", "Japanese", "European"))
# fit1 <- lm(mpg~origin , data = data1)
# summary(fit1)
# data1$origin <- factor(data1$origin, levels = c("Japanese", "American", "European"))
# fit2 <- lm(mpg~origin , data = data1)
# summary(fit2)

# There are notable differences between American and European cars as well as American and Japanese cars and a smaller but still significant difference between European and Japanese cars.# 

data1$origin <- factor(data1$origin, levels = c("American", "Japanese", "European"))
fit1 <- lm(mpg~origin , data = data1)
Anova(fit1)

```
b) Describe on average which origin has the highest mpg and what it is. Which origin has the smallest mpg on average and what is it?

On average, cars of Japanese origin have the highest MPG with 30.5 and American cars have the lowest average MPG with 20.0.

``` {r group_summ}
data1 %>%group_by(origin) %>%summarise(mean_mpg = mean(mpg),sd  = sd(mpg),n = n())
```

c) Are Asian cars more efficient than American cars? Produce a 95% CI's for the difference. 

The difference is -10.417 with a confidence interval ranging from -12.073 to -8.761. Thus Asian cars are more efficient than American cars as 0 is not in the confidence interval.

``` {r}
data1$origin <- factor(data1$origin, levels = c("Japanese", "American", "European"))
summary(lm(mpg ~ origin, data = data1))

```

ii. Try to build a final model which includes `year` and `origin`. 
``` {r build_model}
# Using alpha = 1%
Anova(lm(mpg ~ cylinders + displacement + horsepower + weight + acceleration + year + origin, data = data1))
# Remove accel as it is least relevant
Anova(lm(mpg ~ cylinders + displacement + horsepower + weight + year + origin, data = data1))
# Remove displacement
Anova(lm(mpg ~ cylinders + horsepower + weight + year + origin, data = data1))
# Everything else is relevant at this point
model_final <- lm(mpg ~ cylinders + horsepower + weight + year + origin, data = data1)

```

a) Describe the final model. Include diagnostic plots with particular focus on the model residuals and diagnoses.

The final model is as such. Given the base cylinder count is 2, we know that 4 cylinder cars have 7.67 MPG more on average, 5 cylinder cars have 8.39 more MPG on average, 6 cylinder cars have 5.24 more MPG on average, and 8 cylinder cars have 8.03 more MPG on average. Increasing horsepower while holding all else constant decreases MPG by -0.0254 per horsepower. Increasing the weight by one pound while holding all other parameters constant decreases the MPG by -0.00510. Cars made one year later with all else being held constant have 0.0722 more MPG on average. American cars have 2.21 less MPG on average than Japanese cars, but there is no significant difference between Japanese cars and European cars.

``` {r diag_plots}

summary(model_final)

par(mfrow=c(1,2), mar=c(5,2,4,2), mgp=c(3,0.5,0))
plot(model_final, 1, pch=16) # residual plot. try pch=1 to 25
abline(h=0, col="blue", lwd=3)
# Do some confindence interval bars
abline(h=2*3.12, col="red", lwd=1)
abline(h=-2*3.12, col="red", lwd=1)
plot(model_final, 2) # qqplot
```

b) Describe the year effect and origin effect in this final model.
Are cars being built more efficiently over time? Are Asian cars more efficient than cars built in America or Europe?

Cars are being built more efficiently over time. Asian cars are more efficient than American cars but not European cars.


## Prediction

Use the final model to predict the `mpg` of the following car: A red car built in the US in 1983 that is 180 inches long, has eight cylinders, displaces 350 cu. inches, weighs 4000 pounds, and has a horsepower of 260. Also give a 95% CI for your prediction.

The predicted value is 19.5 with a Ci of (12.9, 26.1)

``` {r final}

length <- c(180)
origin <- c("American")
year <- c(83)
cylinders <- as.factor(c(8))
displacement <- 350
weight <- 4000
horsepower <- 260

predict(model_final, newdata = data.frame(cylinders, horsepower, weight, year, origin), interval = "prediction")

```






