---
title: "Environmental Legislation and its Effects on Air Quality"
author:
- Adithya Ayanam
- Alexander Chen
- Andrew Oliver
- Kamron Ramelmeier
- Sahej Singh
- Raymond Xu
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
    number_sections: yes
  pdf_document:
    toc_depth: '4'
    number_sections: yes
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.width=8, fig.height=4,
                      warning = F, width = 60)
options(scipen = 0, digits = 3)

if(!require('pacman')) install.packages('pacman')

pacman::p_load(Rcpp, geofacet, tidyverse, dplyr, ggplot2, data.table, lubridate, plotROC, usmap, glmnet, 
               plotly, haven, magrittr, Stat2Data, skimr, GGally, rvest, gridExtra, scales, usdata, party, car, randomForest)
```

## Introduction

The goal of this study is to examine the impact of certain variables on the climate 
by examining the AQI of counties across the United States of America using data collected
by the EPA.

There are two smaller sub studies in this presentation: One examining the effects of
the Climate Alliance legislative program, and another examining the correlation between
aspects of counties and the air quality.



## Reading the Data and EDA

To begin we read the data in from the EPA datasets.

\tiny
``` {r readin, results = F, results = 'hide'}
data2020 <- fread("data/annual_aqi_by_county_2020.csv", stringsAsFactors = TRUE)
data2019 <- fread("data/annual_aqi_by_county_2019.csv", stringsAsFactors = TRUE)
data2018 <- fread("data/annual_aqi_by_county_2018.csv", stringsAsFactors = TRUE)
data2017 <- fread("data/annual_aqi_by_county_2017.csv", stringsAsFactors = TRUE)
data2016 <- fread("data/annual_aqi_by_county_2016.csv", stringsAsFactors = TRUE)
data2015 <- fread("data/annual_aqi_by_county_2015.csv", stringsAsFactors = TRUE)
data2014 <- fread("data/annual_aqi_by_county_2014.csv", stringsAsFactors = TRUE)
data2013 <- fread("data/annual_aqi_by_county_2013.csv", stringsAsFactors = TRUE)
data2012 <- fread("data/annual_aqi_by_county_2012.csv", stringsAsFactors = TRUE)
data2011 <- fread("data/annual_aqi_by_county_2011.csv", stringsAsFactors = TRUE)
data2010 <- fread("data/annual_aqi_by_county_2010.csv", stringsAsFactors = TRUE)

raw_data <- rbind(data2020, data2019, data2018, data2017, data2016,
                  data2015, data2014, data2013, data2012, data2011, data2010)

raw_data <- raw_data %>% rename(
  num.days = `Days with AQI`,
  good.days = `Good Days`,
  mod.days = `Moderate Days`,
  sens.group.days = `Unhealthy for Sensitive Groups Days`,
  unhlthy.days = `Unhealthy Days`,
  very.unhlthy.days = `Very Unhealthy Days`,
  haz.days = `Hazardous Days`,
  max.aqi = `Max AQI`,
  med.aqi = `Median AQI`,
  state = State
)

raw_data <- raw_data[raw_data$state != "Country Of Mexico",]
raw_data <- raw_data[raw_data$state != "Puerto Rico",]
raw_data <- raw_data[raw_data$state != "Virgin Islands",]
raw_data <- raw_data[raw_data$state != "District Of Columbia",]
raw_data <- droplevels(raw_data)



raw_data <- raw_data %>%
  mutate(good.pct = good.days/num.days*100,
         mod.pct = mod.days/num.days*100,
         sens.group.pct = sens.group.days/num.days*100,
         unhlthy.pct = unhlthy.days/num.days*100,
         very.unhlthy.pct = very.unhlthy.days/num.days*100,
         haz.pct = haz.days/num.days*100)

mean.state.df <- raw_data %>% group_by(state, Year) %>% summarize(mean.state = mean(med.aqi),
                                                                  peak.state = max(max.aqi))

mean.state.df <- mean.state.df %>% group_by(state) %>%
  mutate(delta.aqi.state = mean.state - lag(mean.state))

mean.state.df$is.climate.alli <- "no"

mean.state.df[mean.state.df$state %in% c("California", "Colorado", "Connecticut", "Delaware",
                               "Hawaii", "Illinois", "Louisiana", "Maine", "Maryland",
                               "Massachucets", "Michigan", "Minnesota", "Nevada",
                               "New Jersey", "New Mexico", "New York", "North Carolina",
                               "Oregon", "Pennsylvania", "Rhode Island",
                               "Vermont", "Virginia", "Washington", "Wisconsin"
), c("is.climate.alli")] <- "yes"

mean.state.df$is.climate.alli.early <- "no"

mean.state.df[mean.state.df$state %in% c("California", "Colorado", "Connecticut", "Delaware",
                               "Hawaii", "Louisiana",
                               "Massachucets","Minnesota",
                               "New York",
                               "Oregon", "Rhode Island",
                               "Vermont", "Virginia", "Washington", "Wisconsin"
), c("is.climate.alli.early")] <- "yes"

mean.state.df$is.climate.alli <- as.factor(mean.state.df$is.climate.alli)

mean.state.df <- mean.state.df %>% mutate(log.aqi = log(mean.state),
                                      log.delta = log(delta.aqi.state))


raw_data <- merge(raw_data, mean.state.df, by=c("state", "Year"))

subbed_data <- raw_data %>% select(state, Year, County, num.days, max.aqi, med.aqi,
                                   good.pct, mod.pct, sens.group.pct, unhlthy.pct,
                                   very.unhlthy.pct, haz.pct, mean.state, is.climate.alli, delta.aqi.state)


summary(lm(data = subbed_data, med.aqi ~ Year))



```

To get a sense of the AQI of US states over the ten year period, we plot a heatmap
by time.

```{r heatmap}
codes <- as.data.frame(cbind(state.abb, state.name)) %>% rename("state" = state.name)
mean.state.df <- merge(mean.state.df, codes, by = "state")

max(mean.state.df$mean.state)

usAqi.plot <- plot_geo(mean.state.df,
                  locationmode = 'USA-states',
                  frame = ~Year) %>% add_trace(locations = ~state.abb,
                                               z = ~mean.state,
                                               color = ~mean.state,
                                               colorscale = 'brwnyl',
                                               zmin = 0, 
                                               zmax = 100) %>%
  layout(geo = list(scope = 'usa'),
         title = "[2010:2020] Avg Air Quality Index by State") %>%
  config(displayModeBar = F)
usAqi.plot
```

---

A more accurate graphical representation by state can be obtained by plotting the AQI by year
over several graphs of state.

``` {r facetAqi, fig.height = 8}
raw_data %>%
  ggplot(aes(x = Year, y = mean.state, group=state,color = state)) +
  geom_line() +
  facet_wrap(~state) +
  theme_bw() +
  theme(legend.position = 0) +
  ggtitle("`State`vs`Avg AQI")
```

The 6 most dangerous pollutants are ozone, nitrogen dioxide, sulfur dioxide, lead, carbon monoxide, and particulate matter.
```{r pollutants}
#EDA 
airquality_ozone<- read.csv("data/OzoneNational.csv", header=T)
airquality_nitrogen<- read.csv("data/Nitrogen_DioxideNational.csv", header=T)
airquality_sulfur<-read.csv("data/Sulfur_DioxideNational.csv", header=T)
airquality_lead<-read.csv("data/LeadNational.csv", header=T)
airquality_carbon<-read.csv("data/Carbon_MonoxideNational.csv", header=T)
airquality_PM10<-read.csv("data/PM10National.csv", header=T)
airquality_PM25<-read.csv("data/PM25National.csv", header=T)
```
```{r Nitrogen Dioxide National}
airquality_nitrogen %>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Nitrogen") +
  xlab("Year") +
  ggtitle("Nitrogen Concentration by year")
```
```{r Ozone National}
airquality_ozone %>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") +
  xlab("Year") +
  ggtitle("Ozone Concentration by year")
```
```{r Sulphur dioxide National}
airquality_sulfur %>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Sulfur Dioxide") +
  xlab("Year") +
  ggtitle("Sulfur Dioxide Concentration by year")
```
```{r Lead National}
airquality_lead %>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Lead") +
  xlab("Year") +
  ggtitle("Lead Concentration by year")
```
```{r Carbon Monoxide National}
airquality_carbon%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Carbon Monoxide") +
  xlab("Year") +
  ggtitle("Carbon Monoxide Concentration by year")
```
```{r PM_10 National}
airquality_PM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("Particulate Matter 10 Concentration by year")
```
```{r PM_25 National}
airquality_PM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("Particulate Matter 25 Concentration by year")
```
These are the graphs of the pollutants over year. As shown in the graphs, the concentrations of each pollutant have decreased due to effective environmental regulation. The graphs show the 10th percentile, 90th percentile, and the mean concentrations of the pollutants.


The plots show that the concentrations have gradually decreased over time for this pollutant or in a few cases have remained the same. 

Much of this improvement can be attributed to the Clean Air Act and its amendments in
1977 and 1990. 

The 2011 reduction in lead concentration might be a result of changes to the Safe
Drinking Water Act.

Regional Graphs
#EDA
#Here are the graphs of the 6 most dangerous pollutants' levels over the past 20 years. 
 ``` {r Region}
airquality_centralozone<-read.csv("data/OzoneCentral.csv", header=T)
airquality_UpperMidwestozone<-read.csv("data/OzoneUpperMidwest.csv", header=T)
airquality_Northeastozone<-read.csv("data/OzoneNortheast.csv", header=T)
airquality_Northwestozone<-read.csv("data/OzoneNorthwest.csv", header=T)
airquality_Southozone<-read.csv("data/OzoneSouth.csv", header=T)
airquality_Southeastozone<-read.csv("data/OzoneSoutheast.csv", header=T)
airquality_Southwestozone<-read.csv("data/OzoneSouthwest.csv", header=T)
airquality_Westozone<-read.csv("data/OzoneWest.csv", header=T)
airquality_NRPozone<-read.csv("data/OzoneNorthernRockies.csv", header=T)

p1<-airquality_centralozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone Central")
 
  p2 <- airquality_UpperMidwestozone%>% #upper midwest
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone MW")
  
  p3 <- airquality_Northeastozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone NE")
  
  p4 <- airquality_Northwestozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone NW")
  
  p5 <- airquality_Southozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone South")
  
  p6 <- airquality_Southeastozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone SE")
    
  p7 <- airquality_Southwestozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone SW")
     
  p8 <- airquality_Westozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone West")
 
  p9 <- airquality_NRPozone%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone NRP") #North Rockie and Plains
 
grid.arrange(p1, p2, p3,p4,p5,p6,p7,p8,p9,  ncol=3)# facet the two plots  side by side
# The average levels of ozone has decreased visibly in most places except the Western part of the United States, which includes the Southwest, the West, and the Northern Rockies and Plains where the ozone levels are basically the same. The Western states receive plenty of sunlight and the cities with the most amount of sunlight come from the West. This plays a major role in the high levels of ozone. Also Trans Pacific Pollution plays a major role in the level of ozone. 
```
```{r Region wise Nitrogen Di oxide }
airquality_centralnitrogen<-read.csv("data/Nitrogen_DioxideCentral.csv", header=T)
airquality_UpperMidwestnitrogen<-read.csv("data/Nitrogen_DioxideUpperMidwest.csv", header=T)
airquality_Northeastnitrogen<-read.csv("data/Nitrogen_DioxideNortheast.csv", header=T)
airquality_Southnitrogen<-read.csv("data/Nitrogen_DioxideSouth.csv", header=T)
airquality_Southeastnitrogen<-read.csv("data/Nitrogen_DioxideSoutheast.csv", header=T)
airquality_Southwestnitrogen<-read.csv("data/Nitrogen_DioxideSouthwest.csv", header=T)
airquality_Westnitrogen<-read.csv("data/Nitrogen_DioxideWest.csv", header=T)
airquality_NRPnitrogen<-read.csv("data/Nitrogen_DioxideNorthernRockies.csv", header=T)

p11<-airquality_centralnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 Central")

p12 <- airquality_UpperMidwestnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 MW")

p13 <- airquality_Northeastnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 NE")

 

p14 <- airquality_Southnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 South")

p15 <- airquality_Southeastnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 SE")



p16 <- airquality_Southwestnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 SW")

p17 <- airquality_Westnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 West")

 

p18 <- airquality_NRPnitrogen%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 NRP")

grid.arrange(p11, p12, p13,p14,p15,p16,p17,p18,  ncol=3)# facet the two plots  side by side

# The average levels of nitrogen dioxide have been decreasing over the past 20 years throughout the US. Due to the rural nature, the Northern Rockies and Plains had the least amount of nitrogen dioxide while regions that contained more urban areas such as the Central Region, which contained Chicago and the Northeast region, which contained New York had more nitrogen dioxide to begin with. Places that encompassed a wide range of rural and urban locations had very extreme values for the 10th and the 90th percentile relative to the mean and places such as Northern Rockies and Plains which was generally rural throughout had a 10th and 90th percentile that were not far away from the mean. 
```

```{r Regionwise Sulphurdioxide}
airquality_centralsulfurdioxide<-read.csv("data/Sulfur_DioxideCentral.csv", header=T)
airquality_UpperMidwestsulfurdioxide<-read.csv("data/Sulfur_DioxideUpperMidwest.csv", header=T)
airquality_Northeastsulfurdioxide<-read.csv("data/Sulfur_DioxideNortheast.csv", header=T)
airquality_Southsulfurdioxide<-read.csv("data/Sulfur_DioxideSouth.csv", header=T)
airquality_Southeastsulfurdioxide<-read.csv("data/Sulfur_DioxideSoutheast.csv", header=T)
airquality_Southwestsulfurdioxide<-read.csv("data/Sulfur_DioxideSouthwest.csv", header=T)
airquality_Westsulfurdioxide<-read.csv("data/Sulfur_DioxideWest.csv", header=T)
airquality_NRPsulfurdioxide<-read.csv("data/Sulfur_DioxideNorthernRockies.csv", header=T)



p21<-airquality_centralsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 Central")

p22 <- airquality_UpperMidwestsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 MW")

p23 <- airquality_Northeastsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 NE")

 

p24 <- airquality_Southsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 South")

p25 <- airquality_Southeastsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 SE")



p26 <- airquality_Southwestsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 SW")

p27 <- airquality_Westsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 West")

 

p28 <- airquality_NRPsulfurdioxide%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 NRP")

grid.arrange(p21, p22, p23,p24,p25,p26,p27,p28,  ncol=3)# facet the two plots  side by side
#Throughout the country the average concentration of SO2 has decreased substantially over the past 20 years with the Central Region's average concentration of SO2 decreasing the most. Due to coal fired powerplants, the Central Region had the highest concentration of SO2 when the data collection started and over the past 20 years, due to regulations and technologies that promoted cleaner burning, the SO2 concentrations have declined sharply in that region.  
```
```{r Regionwise CO}
airquality_centralcarbonmonoxide<-read.csv("data/Carbon_MonoxideCentral.csv", header=T)
airquality_UpperMidwestcarbonmonoxide<-read.csv("data/Carbon_MonoxideUpperMidwest.csv", header=T)
airquality_Northeastcarbonmonoxide<-read.csv("data/Carbon_MonoxideNortheast.csv", header=T)
airquality_Northwestcarbonmonoxide<-read.csv("data/Carbon_MonoxideNorthwest.csv", header=T)
airquality_Southcarbonmonoxide<-read.csv("data/Carbon_MonoxideSouth.csv", header=T)
airquality_Southeastcarbonmonoxide<-read.csv("data/Carbon_MonoxideSoutheast.csv", header=T)
airquality_Southwestcarbonmonoxide<-read.csv("data/Carbon_MonoxideSouthwest.csv", header=T)
airquality_Westcarbonmonoxide<-read.csv("data/Carbon_MonoxideWest.csv", header=T)
 

p41<-airquality_centralcarbonmonoxide%>% 
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO Central")

p42 <- airquality_UpperMidwestcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO MW") #Midwest

p43 <- airquality_Northeastcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO NE")

p44 <- airquality_Northwestcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO NW")

p45 <- airquality_Southcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO South")

p46 <- airquality_Southeastcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO SE")

p47 <- airquality_Southwestcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO SW")

p48 <- airquality_Westcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO West")

grid.arrange(p41, p42, p43,p44,p45,p46,p47,p48, ncol=3)# facet the two plots  side by side
In every region except the Northwest, carbon monoxide has been decreasing over the past 20 years with the South and the Southwest having a marked decline in average concentration of CO. The regions with the most Carbon monoxide are areas with higher latitude and higher altitude. 
```
```{r Regionwise PM10 }
airquality_centralPM10<-read.csv("data/PM10Central.csv", header=T)
airquality_UpperMidwestPM10<-read.csv("data/PM10UpperMidwest.csv", header=T)
airquality_NortheastPM10<-read.csv("data/PM10Northeast.csv", header=T)
airquality_NorthwestPM10<-read.csv("data/PM10Northwest.csv", header=T)
airquality_SouthPM10<-read.csv("data/PM10South.csv", header=T)
airquality_SoutheastPM10<-read.csv("data/PM10Southeast.csv", header=T)
airquality_SouthwestPM10<-read.csv("data/PM10Southwest.csv", header=T)
airquality_WestPM10<-read.csv("data/PM10West.csv", header=T)
airquality_NRPPM10<-read.csv("data/PM10NorthernRockies.csv", header=T)

p51<-airquality_centralPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 Central")

p52 <- airquality_UpperMidwestPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 MW")

p53 <- airquality_NortheastPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 NE")

p54 <- airquality_NorthwestPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 NW")

p55 <- airquality_SouthPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 South")

p56 <- airquality_SoutheastPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 SE")



p57 <- airquality_SouthwestPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 SW")

p58 <- airquality_WestPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 West")
 

p59 <- airquality_NRPPM10%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 NRP")

grid.arrange(p51, p52, p53,p54,p55,p56,p57,p58,p59, ncol=3)# facet the two plots  side by side
# Average PM10 concentration has significantly decreased in the West, increased in the Northwest, and did not change much in the rest of the regions. 

```{r Regionwise PM25}
airquality_centralPM25<-read.csv("data/PM25Central.csv", header=T)
airquality_UpperMidwestPM25<-read.csv("data/PM25UpperMidwest.csv", header=T)
airquality_NortheastPM25<-read.csv("data/PM25Northeast.csv", header=T)
airquality_NorthwestPM25<-read.csv("data/PM25Northwest.csv", header=T)
airquality_SouthPM25<-read.csv("data/PM25South.csv", header=T)
airquality_SoutheastPM25<-read.csv("data/PM25Southeast.csv", header=T)
airquality_SouthwestPM25<-read.csv("data/PM25Southwest.csv", header=T)
airquality_WestPM25<-read.csv("data/PM25West.csv", header=T)
airquality_NRPPM25<-read.csv("data/PM25NorthernRockies.csv", header=T)

p61<-airquality_centralPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 Central")

p62 <- airquality_UpperMidwestPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 MW")

p63 <- airquality_NortheastPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 NE")

p64 <- airquality_NorthwestPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 NW")

p65 <- airquality_SouthPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 South")

p66 <- airquality_SoutheastPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 SE")



p67 <- airquality_SouthwestPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 SW")

p68 <- airquality_WestPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 West")
 
p69 <- airquality_NRPPM25%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 NRP")

grid.arrange(p61, p62, p63,p64,p65,p66,p67,p68,p69, ncol=3)# facet the two plots  side by side
#In every region except the Western regions of the US, Particulate matter has declined which significant decline in average concentration of PM 25 in the central region. The Northwest has had an increase in Particulate Matter concentration and the West and the Southwest a regions have not have their average concentrations change much and in the past few years, due to wild fires, particulate matter has actually increased in the West and the Southwest. 
```

## Climate Alliance

The Climate Alliance is a bipartisan group of governors who aim to reduce their emissions.
Formed in response to the US withdrawal from the Paris Accord, it acts on the state level.

To determine the effectiveness of the Climate Alliance, we look if there is a significant
difference in their rate of change of AQI within Climate Alliance states. A significant difference would indicate that the Climate Alliance was effective
in improving the AQI of its. states. We look at data from 2019 (so thus the change in AQI
from 2018 to 2019) to observe the effect after the formation.

``` {r CA Linear Regression}
mean.state.df$is.climate.alli <- as.factor(mean.state.df$is.climate.alli)

summary(lm(data = mean.state.df %>% filter(Year == 2019), delta.aqi.state ~ is.climate.alli))
```

While on average Climate Alliance states do improve faster, the difference is not significant at a 
0.05 level. Despite this, there is a notable change in the average rate of change.

It is possible that the data we are using also might be obscuring some of the effects as 
not all states joined in 2017. The last state to join, Montana, joined in mid 2019. States 
like those make the impact of the Climate Alliance hard to see.

To be more precise, we look at only states who joined within the first month.

``` {r CA Linear Regression Early}
mean.state.df$is.climate.alli <- as.factor(mean.state.df$is.climate.alli.early)

summary(lm(data = mean.state.df %>% filter(Year == 2019), delta.aqi.state ~ is.climate.alli.early))
```
The difference here is larger and more significant but is still insignificant at the 0.05 level.

While our study is does not conclude that the Climate Alliance is definitely improving the air
quality of its member states, its members are improving slightly faster on average at a 75% confidence level.

## County Level Effects on AQI

Using the data found by the USDA’s Economic Research Service, we look for predictors in counties to determine air quality and find correlations.

This begins by merging the 2019 AQI with the latest USDA ERS data. We use 2019 data to avoid skewing due to the 2020 West Coast fires.

To begin the analysis, we start by merging county data with AQI data. We start by
merging all three sets of ERS county data, and then we merge by county and state.  

Break the cleaned and merged dataset into X and Y for use with cv.glmnet. We use
set.seed(1) for consistency.

Looking at the histogram shows that the county AQI data appears to be normal.


``` {r cleanmerge}
county_data_jobs <- fread("data/Release23_June2021/Jobs.csv")
county_data_people <- fread("data/Release23_June2021/People.csv")
county_data_income <- fread("data/Release23_June2021/Income.csv")


county_data_master <- merge(county_data_jobs, county_data_people, by = c("County", "FIPS", "State"))
county_data_master <- merge(county_data_master, county_data_income, by = c("County", "FIPS", "State"))
county_data_master$State <- abbr2state(county_data_master$State)
county_data_master <- county_data_master %>% rename(state = State)

dated_subbed_data <- subbed_data[subbed_data$Year == 2019,] %>% select(state, Year, County, med.aqi, delta.aqi.state)

master_data <- merge(county_data_master, dated_subbed_data, by = c("County", "state"))

glmnet_data <- master_data %>% select(-County, -FIPS) %>% drop_na()

hist(glmnet_data$med.aqi)
```

\tiny
``` {r glmnet}

set.seed(1)
X <- model.matrix(delta.aqi.state~.-state, glmnet_data)[,-1]
y <- glmnet_data$delta.aqi.state


fit.cv <- cv.glmnet(X, y, alpha = 1, nfolds = 10)

coeffs <- coef(fit.cv)
nonzeros <- rownames(as.matrix(coeffs[which(coeffs != 0), 0]))
non_state_intercept_coef <- nonzeros[2:length(nonzeros)]

select_cols <- c("med.aqi", non_state_intercept_coef)
lm_cols <- glmnet_data %>% select(select_cols)

fit.lm <- lm(data = lm_cols, med.aqi ~ .)

Anova(fit.lm)

```




We remove the variables that are the least relevant to see what factors remain.
\tiny
``` {r anova1, include = FALSE}
fit.lm <- update(fit.lm, .~. -PctEmpMining)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -ForeignBornCaribPct)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -Ed3SomeCollegePct)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpTrans)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -AsianNonHispanicPct2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpManufacturing)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -LandAreaSQMiles2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -UnempRate2012)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PopChangeRate1819)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -UnempRate2020)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -MultipleRacePopChangeRate0010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpAgriculture)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -Immigration_Rate_2000_2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -ForeignBornEuropePct)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpChange1920)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -MultipleRacePct2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -Net_International_Migration_Rate_2010_2019)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -Deep_Pov_All)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpConstruction)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -NetMigrationRate0010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -NaturalChangeRate0010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -NaturalChangeRate0010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -NaturalChangeRate1010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -NaturalChangeRate1019)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -ForeignBornMexPct)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -BlackNonHispanicPct2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -NativeAmericanNonHispanicPct2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -ForeignBornAfricaNum)
Anova(fit.lm)
```


```{r assign final}
fit.final <- fit.lm
summary(fit.lm)
```




From the final model, we see that most of the impact on AQI is geographical. For example, 
the increase from ForeignBornMexNum could signal that states closer to
the Mexican border tend to have worse AQIs due to their location. However, the
most clear predictors are the states themselves.

The assumptions for linearity appear to hold up until about 1 standard deviation below
the mean.

``` {r plots, fig.show="hold", out.width="50%"}
plot(fit.final, 1)
plot(fit.final, 2)
```


```{r CA Random Forest, eval = F}
set.seed(467)
ind <- sample(2, nrow(lm_cols), replace = T, prob = c(0.7, 0.3))
train <- lm_cols[ind==1,]
test <- lm_cols[ind==2,]

library(randomForest)
set.seed(200)
rf.fit <- randomForest(med.aqi~., data = train, ntree = 500)

plot(rf.fit, col="red", pch=16, type="p", 
     main="default plot, ")

rf.error.p <- 1:19
for (i in 1:19)
{
  rf.fit <- randomForest(med.aqi~., train, mtry=i, ntree=500)
  rf.error.p[i] <- rf.fit$mse[500]
}

plot(1:19, rf.error.p, pch=16,
     main = "Testing Errors: ntree = 500",
     xlab="mtry",
     ylab="OOB MSE of mtry")
lines(1:19, rf.error.p)

set.seed(250)
rf.fit <- randomForest(med.aqi~., train, mtry=2, ntree=500)
rf.fit.pred <- predict(rf.fit, train)
MSE_test2 <- mean((train$med.aqi - rf.fit.pred)^2)
MSE_test3 <- mean((test$med.aqi - rf.fit.pred)^2)

rf.fit$importance
attributes(rf.fit)

varImpPlot(rf.fit, sort=TRUE, n.var=min(30, nrow(rf.fit$importance)))

plot(rf.fit)
```





## Conclusion

The overall objective of this study was to use the AQI of counties across the USA to determine the impact of variables on the climate. Using data collected by the EPA, we were able to focus on the effect of the Climate Alliance on curbing the deterioration of the AQI across the nation, as well as the correlation between aspects of counties and their air quality.

From this study, we were able to conclude that the Climate Alliance has not had much of an effect yet on the AQI of member states, but do have better AQIs on average compared to other states. We were also able to see that most of the impact on the AQI is geographical based on the significant variables of the model. 

