---
title: "Environmental Legislation and its Effects on Air Quality"
author:
- Adithya Ayanam
- Alexander Chen
- Andrew Oliver
- Kamron Ramelmeier
- Sahej Singh
- Raymond Xu
output: 
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.width=8, fig.height=4,
                      warning = F, width = 60)
options(scipen = 0, digits = 3)

if(!require('pacman')) install.packages('pacman')

pacman::p_load(Rcpp, geofacet, tidyverse, dplyr, ggplot2, data.table, lubridate, plotROC, usmap, glmnet, 
               plotly, haven, magrittr, Stat2Data, skimr, GGally, rvest, gridExtra, scales, usdata, party, car, randomForest)
```

## Introduction

The goal of this study is to examine the impact of certain variables on the climate 
by examining the AQI of counties across the United States of America using data collected
by the EPA.

There are two smaller sub studies in this presentation: One examining the effects of
the Climate Alliance legislative program, and another examining the correlation between
aspects of counties and the air quality.



## Reading the Data and EDA


\frametitle{Read and Clean}
To begin we read the data in from the EPA datasets.

\tiny
``` {r readin, results = F, results = 'hide'}
data2020 <- fread("data/annual_aqi_by_county_2020.csv", stringsAsFactors = TRUE)
data2019 <- fread("data/annual_aqi_by_county_2019.csv", stringsAsFactors = TRUE)
data2018 <- fread("data/annual_aqi_by_county_2018.csv", stringsAsFactors = TRUE)
data2017 <- fread("data/annual_aqi_by_county_2017.csv", stringsAsFactors = TRUE)
data2016 <- fread("data/annual_aqi_by_county_2016.csv", stringsAsFactors = TRUE)
data2015 <- fread("data/annual_aqi_by_county_2015.csv", stringsAsFactors = TRUE)
data2014 <- fread("data/annual_aqi_by_county_2014.csv", stringsAsFactors = TRUE)
data2013 <- fread("data/annual_aqi_by_county_2013.csv", stringsAsFactors = TRUE)
data2012 <- fread("data/annual_aqi_by_county_2012.csv", stringsAsFactors = TRUE)
data2011 <- fread("data/annual_aqi_by_county_2011.csv", stringsAsFactors = TRUE)
data2010 <- fread("data/annual_aqi_by_county_2010.csv", stringsAsFactors = TRUE)

raw_data <- rbind(data2020, data2019, data2018, data2017, data2016,
                  data2015, data2014, data2013, data2012, data2011, data2010)

raw_data <- raw_data %>% rename(
  num.days = `Days with AQI`,
  good.days = `Good Days`,
  mod.days = `Moderate Days`,
  sens.group.days = `Unhealthy for Sensitive Groups Days`,
  unhlthy.days = `Unhealthy Days`,
  very.unhlthy.days = `Very Unhealthy Days`,
  haz.days = `Hazardous Days`,
  max.aqi = `Max AQI`,
  med.aqi = `Median AQI`,
  state = State
)

raw_data <- raw_data[raw_data$state != "Country Of Mexico",]
raw_data <- raw_data[raw_data$state != "Puerto Rico",]
raw_data <- raw_data[raw_data$state != "Virgin Islands",]
raw_data <- raw_data[raw_data$state != "District Of Columbia",]
raw_data <- droplevels(raw_data)



raw_data <- raw_data %>%
  mutate(good.pct = good.days/num.days*100,
         mod.pct = mod.days/num.days*100,
         sens.group.pct = sens.group.days/num.days*100,
         unhlthy.pct = unhlthy.days/num.days*100,
         very.unhlthy.pct = very.unhlthy.days/num.days*100,
         haz.pct = haz.days/num.days*100)

mean.state.df <- raw_data %>% group_by(state, Year) %>% summarize(mean.state = mean(med.aqi),
                                                                  peak.state = max(max.aqi))

mean.state.df <- mean.state.df %>% group_by(state) %>%
  mutate(delta.aqi.state = mean.state - lag(mean.state))

mean.state.df$is.climate.alli <- "no"

mean.state.df[mean.state.df$state %in% c("California", "Colorado", "Connecticut", "Delaware",
                               "Hawaii", "Illinois", "Louisiana", "Maine", "Maryland",
                               "Massachucets", "Michigan", "Minnesota", "Nevada",
                               "New Jersey", "New Mexico", "New York", "North Carolina",
                               "Oregon", "Pennsylvania", "Rhode Island",
                               "Vermont", "Virginia", "Washington", "Wisconsin"
), c("is.climate.alli")] <- "yes"

mean.state.df$is.climate.alli <- as.factor(mean.state.df$is.climate.alli)

raw_data <- merge(raw_data, mean.state.df, by=c("state", "Year"))

subbed_data <- raw_data %>% select(state, Year, County, num.days, max.aqi, med.aqi,
                                   good.pct, mod.pct, sens.group.pct, unhlthy.pct,
                                   very.unhlthy.pct, haz.pct, mean.state, is.climate.alli, delta.aqi.state)

summary(lm(data = subbed_data, med.aqi ~ Year))



```

\frametitle{Heatmap}

```{r heatmap}
codes <- as.data.frame(cbind(state.abb, state.name)) %>% rename("state" = state.name)
mean.state.df <- merge(mean.state.df, codes, by = "state")

max(mean.state.df$mean.state)

usAqi.plot <- plot_geo(mean.state.df,
                  locationmode = 'USA-states',
                  frame = ~Year) %>% add_trace(locations = ~state.abb,
                                               z = ~mean.state,
                                               color = ~mean.state,
                                               colorscale = 'brwnyl',
                                               zmin = 0, 
                                               zmax = 100) %>%
  layout(geo = list(scope = 'usa'),
         title = "[2010:2020] Avg Air Quality Index by State") %>%
  config(displayModeBar = F)
usAqi.plot
```

---
\frametitle{AQI by Year Faceted by State}

``` {r facetAqi}
raw_data %>%
  ggplot(aes(x = Year, y = mean.state, group=state,color = state)) +
  geom_line() +
  facet_wrap(~state) +
  theme_bw() +
  theme(legend.position = 0) +
  ggtitle("`State`vs`Avg AQI")
```
\frametitle{6 most dangerous Pollutants }
The 6 most dangerous pollutants are ozone, nitrogen dioxide, sulfur dioxide, lead, carbon monoxide, and particulate matter.
```{r pollutants, eval = F}
#EDA 
airquality_ozone<- read.csv("data/OzoneNational.csv", header=T)
airquality_nitrogen<- read.csv("data/Nitrogen_DioxideNational.csv", header=T)
airquality_sulfur<-read.csv("data/Sulfur_DioxideNational.csv", header=T)
airquality_lead<-read.csv("data/LeadNational.csv", header=T)
airquality_carbon<-read.csv("data/Carbon_MonoxideNational.csv", header=T)
airquality_PM10<-read.csv("data/PM10National.csv", header=T)
airquality_PM25<-read.csv("data/PM25National.csv", header=T)
```
```{r Nitrogen Di Oxide National}
str( airquality_nitrogen)# data format
summary( airquality_nitrogen)# quick summary. missing values may be shown
airquality_nitrogen %>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Nitrogen") +
  xlab("Year") +
  ggtitle("Nitrogen Concentration by year")
```
```{r Ozone National}
str( airquality_ozone)# data format
summary( airquality_ozone)# quick summary. missing values may be shown
airquality_ozone %>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") +
  xlab("Year") +
  ggtitle("Ozone Concentration by year")
```
```{r Sulphur di oxide National}
str( airquality_sulfur)# data format
summary( airquality_sulfur)# quick summary. missing values may be shown
airquality_sulfur %>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Sulfur Dioxide") +
  xlab("Year") +
  ggtitle("Sulfur Dioxide Concentration by year")
```
```{r Lead National}
str( airquality_lead)# data format
summary( airquality_lead)# quick summary. missing values may be shown
airquality_lead %>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Lead") +
  xlab("Year") +
  ggtitle("Lead Concentration by year")
```
```{r Carbon Monoxide National}
str( airquality_carbon)# data format
summary( airquality_carbon)# quick summary. missing values may be shown
airquality_carbon%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Carbon Monoxide") +
  xlab("Year") +
  ggtitle("Carbon Monoxide Concentration by year")
```
```{r PM_10 National}
str( airquality_PM10)# data format  
summary( airquality_PM10)# quick summary. missing values may be shown
airquality_PM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("Particulate Matter 10 Concentration by year")
```
```{r PM_25 National}
str( airquality_PM25)# data format  
summary( airquality_PM25)# quick summary. missing values may be shown
airquality_PM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("Particulate Matter 25 Concentration by year")
```
#Displays the graphs of the pollutants over the span of 40 years. As shown in the graphs, the concentrations of each pollutant have decreased since the 1980s due to effective environmnetal regulation. The graphs show the 10th percentile, 90th percentile, and the mean concentrations of the p[ollutant. 
```{r pollutants, eval = F}
#EDA 
airquality_centralozone<-read.csv("data/OzoneCentral.csv", header=T)
airquality_UpperMidwestozone<-read.csv("data/OzoneUpperMidwest.csv", header=T)
airquality_Northeastozone<-read.csv("data/OzoneNortheast.csv", header=T)
airquality_Northwestozone<-read.csv("data/OzoneNorthwest.csv", header=T)
airquality_Southozone<-read.csv("data/OzoneSouth.csv", header=T)
airquality_Southeastozone<-read.csv("data/OzoneSoutheast.csv", header=T)
airquality_Southwestozone<-read.csv("data/OzoneSouthwest.csv", header=T)
airquality_Westozone<-read.csv("data/OzoneWest.csv", header=T)
airquality_NRPozone<-read.csv("data/OzoneNorthernRockies.csv", header=T)
```

```{r Regional Ozone, eval = F}
p1<-airquality_centralozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone Central")

p2 <- airquality_UpperMidwestozone%>% #upper midwest
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone MW")

p3 <- airquality_Northeastozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone NE")

p4 <- airquality_Northwestozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone NW")

p5 <- airquality_Southozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone South")

p6 <- airquality_Southeastozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone SE")

p7 <- airquality_Southwestozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone SW")

p8 <- airquality_Westozone%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone West")

p9 <- airquality_NRPozone%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("Ozone") + 
  xlab("Year") +
  ggtitle("Ozone NRP") #North Rockie and Plains

grid.arrange(p1, p2, p3,p4,p5,p6,p7,p8,p9,  ncol=3)# facet the two plots  side by side
```

Here we display the concentration of ozone trend in each region. The ozone levels have been decreasing in the past 20 years due to earlier effective environmental regulation. The Western States have the most ozone due to oil and gas extraction and are also known for having the most amount of sunlight. 

```{r Regional Nitrogen Di oxide, eval = F}
#EDA
airquality_centralnitrogen<-read.csv("data/Nitrogen_DioxideCentral.csv", header=T)
airquality_UpperMidwestnitrogen<-read.csv("data/Nitrogen_DioxideUpperMidwest.csv", header=T)
airquality_Northeastnitrogen<-read.csv("data/Nitrogen_DioxideNortheast.csv", header=T)
airquality_Southnitrogen<-read.csv("data/Nitrogen_DioxideSouth.csv", header=T)
airquality_Southeastnitrogen<-read.csv("data/Nitrogen_DioxideSoutheast.csv", header=T)
airquality_Southwestnitrogen<-read.csv("data/Nitrogen_DioxideSouthwest.csv", header=T)
airquality_Westnitrogen<-read.csv("data/Nitrogen_DioxideWest.csv", header=T)
airquality_NRPnitrogen<-read.csv("data/Nitrogen_DioxideNorthernRockies.csv", header=T)
```
```{r Regional Nitrogen Di oxide, eval = F}
p11<-airquality_centralnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 Central")

p12 <- airquality_UpperMidwestnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 MW")

p13 <- airquality_Northeastnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 NE")

p14 <- airquality_Southnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 South")

p15 <- airquality_Southeastnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 SE")

p16 <- airquality_Southwestnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 SW")

p17 <- airquality_Westnitrogen%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 West")

p18 <- airquality_NRPnitrogen%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("NO2") + 
  xlab("Year") +
  ggtitle("NO2 NRP")

grid.arrange(p11, p12, p13,p14,p15,p16,p17,p18,  ncol=3)# facet the two plots  side by side
```

The northern rockies and plains not withstanding, NO2 seems to be declining throughout every region.
However, this is because No2 is generally very low in that region. 
The areas where NO2 is naturally very high is in the very urban areas most notably the Northeast and California.

```{r Regional Sulphur Di oxide, eval = F} 
#EDA
airquality_centralsulfurdioxide<-read.csv("data/Sulfur_DioxideCentral.csv", header=T)
airquality_UpperMidwestsulfurdioxide<-read.csv("data/Sulfur_DioxideUpperMidwest.csv", header=T)
airquality_Northeastsulfurdioxide<-read.csv("data/Sulfur_DioxideNortheast.csv", header=T)
airquality_Southsulfurdioxide<-read.csv("data/Sulfur_DioxideSouth.csv", header=T)
airquality_Southeastsulfurdioxide<-read.csv("data/Sulfur_DioxideSoutheast.csv", header=T)
airquality_Southwestsulfurdioxide<-read.csv("data/Sulfur_DioxideSouthwest.csv", header=T)
airquality_Westsulfurdioxide<-read.csv("data/Sulfur_DioxideWest.csv", header=T)
airquality_NRPsulfurdioxide<-read.csv("data/Sulfur_DioxideNorthernRockies.csv", header=T)
```

```{r Regional Sulphur Di oxide} 
p21<-airquality_centralsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 Central")

p22 <- airquality_UpperMidwestsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 MW")

p23 <- airquality_Northeastsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 NE")

p24 <- airquality_Southsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 South")

p25 <- airquality_Southeastsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 SE")

p26 <- airquality_Southwestsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 SW")

p27 <- airquality_Westsulfurdioxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 West")

p28 <- airquality_NRPsulfurdioxide%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("SO2") + 
  xlab("Year") +
  ggtitle("SO2 NRP")

grid.arrange(p21, p22, p23,p24,p25,p26,p27,p28,  ncol=3)# facet the two plots  side by side
```

Like the other two pollutants, the concentration of sulfur dioxide has also decreased remarkably. 
Sulfur Dioxide concentrations have been decreasing throughout the country outside of the Southwest in the past several years.
Sulfur is most prevalent in the Southwest and Central Region while being least prevalent in the West.


```{r Regional Carbon Monoxide, eval = F} 
#EDA
airquality_centralcarbonmonoxide<-read.csv("data/Carbon_MonoxideCentral.csv", header=T)
airquality_UpperMidwestcarbonmonoxide<-read.csv("data/Carbon_MonoxideUpperMidwest.csv", header=T)
airquality_Northeastcarbonmonoxide<-read.csv("data/Carbon_MonoxideNortheast.csv", header=T)
airquality_Northwestcarbonmonoxide<-read.csv("data/Carbon_MonoxideNorthwest.csv", header=T)
airquality_Southcarbonmonoxide<-read.csv("data/Carbon_MonoxideSouth.csv", header=T)
airquality_Southeastcarbonmonoxide<-read.csv("data/Carbon_MonoxideSoutheast.csv", header=T)
airquality_Southwestcarbonmonoxide<-read.csv("data/Carbon_MonoxideSouthwest.csv", header=T)
airquality_Westcarbonmonoxide<-read.csv("data/Carbon_MonoxideWest.csv", header=T)
```
 ```{r Regional Carbon Monoxide} 
p41<-airquality_centralcarbonmonoxide%>% 
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO Central")

p42 <- airquality_UpperMidwestcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO MW") #Midwest

p43 <- airquality_Northeastcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO NE")

p44 <- airquality_Northwestcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO NW")

p45 <- airquality_Southcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO South")

p46 <- airquality_Southeastcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO SE")

p47 <- airquality_Southwestcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO SW")

p48 <- airquality_Westcarbonmonoxide%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("CO") + 
  xlab("Year") +
  ggtitle("CO West")

grid.arrange(p41, p42, p43,p44,p45,p46,p47,p48, ncol=3)# facet the two plots  side by side
```
Carbon monoxide concentration decreases in every location except the Northwest where it increases. 
Carbon monoxide is most prevalent in the Northwest right now but it used to be very prevalent in the Central and Western area. 

 ```{r Regional PM10} 
 #EDA
airquality_centralPM10<-read.csv("data/PM10Central.csv", header=T)
airquality_UpperMidwestPM10<-read.csv("data/PM10UpperMidwest.csv", header=T)
airquality_NortheastPM10<-read.csv("data/PM10Northeast.csv", header=T)
airquality_NorthwestPM10<-read.csv("data/PM10Northwest.csv", header=T)
airquality_SouthPM10<-read.csv("data/PM10South.csv", header=T)
airquality_SoutheastPM10<-read.csv("data/PM10Southeast.csv", header=T)
airquality_SouthwestPM10<-read.csv("data/PM10Southwest.csv", header=T)
airquality_WestPM10<-read.csv("data/PM10West.csv", header=T)
airquality_NRPPM10<-read.csv("data/PM10NorthernRockies.csv", header=T)
```
```{r Regional PM10}
p51<-airquality_centralPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 Central")

p52 <- airquality_UpperMidwestPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 MW")

p53 <- airquality_NortheastPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 NE")

p54 <- airquality_NorthwestPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 NW")

p55 <- airquality_SouthPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 South")

p56 <- airquality_SoutheastPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 SE")

p57 <- airquality_SouthwestPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 SW")

p58 <- airquality_WestPM10%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 West")

p59 <- airquality_NRPPM10%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey60") +
  geom_line(aes(y = Mean)) +
  ylab("PM10") + 
  xlab("Year") +
  ggtitle("PM10 NRP")

grid.arrange(p51, p52, p53,p54,p55,p56,p57,p58,p59, ncol=3)# facet the two plots  side by side
```

In the majority of areas with the exception of the Rockies and Plains, the concentration of PM10 has been decreasing for many years.

```{r Regional PM25} 
 #EDA
airquality_centralPM25<-read.csv("data/PM25Central.csv", header=T)
airquality_UpperMidwestPM25<-read.csv("data/PM25UpperMidwest.csv", header=T)
airquality_NortheastPM25<-read.csv("data/PM25Northeast.csv", header=T)
airquality_NorthwestPM25<-read.csv("data/PM25Northwest.csv", header=T)
airquality_SouthPM25<-read.csv("data/PM25South.csv", header=T)
airquality_SoutheastPM25<-read.csv("data/PM25Southeast.csv", header=T)
airquality_SouthwestPM25<-read.csv("data/PM25Southwest.csv", header=T)
airquality_WestPM25<-read.csv("data/PM25West.csv", header=T)
airquality_NRPPM25<-read.csv("data/PM25NorthernRockies.csv", header=T)
```
```{r Regional PM25} 
p61<-airquality_centralPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 Central")

p62 <- airquality_UpperMidwestPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 MW")

p63 <- airquality_NortheastPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 NE")

p64 <- airquality_NorthwestPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 NW")

p65 <- airquality_SouthPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 South")

p66 <- airquality_SoutheastPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 SE")

p67 <- airquality_SouthwestPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 SW")

p68 <- airquality_WestPM25%>%
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 West")

p69 <- airquality_NRPPM25%>% #Northern Rockies and Plains
  ggplot(aes(x=Year )) +
  geom_ribbon(aes(ymin = X10th.Percentile, ymax = X90th.Percentile), fill = "grey70") +
  geom_line(aes(y = Mean)) +
  ylab("PM25") + 
  xlab("Year") +
  ggtitle("PM25 NRP")

grid.arrange(p61, p62, p63,p64,p65,p66,p67,p68,p69, ncol=3)# facet the two plots  side by side
```

The plots show that the concentrations have gradually decreased over time for this pollutant or in a few cases have remained the same. 
The places where the data varies significantly are the West, Southwest, and the Rockies. 

Climate Alliance
---
\frametitle{Linear Regression}

``` {r CA Linear Regression}
subbed_data$is.climate.alli <- as.factor(subbed_data$is.climate.alli)
focus_data <- subbed_data[subbed_data$Year %in% 2016:2019,]
summary(lm(data = mean.state.df, mean.state ~ is.climate.alli + Year))
summary(lm(data = mean.state.df, delta.aqi.state ~ is.climate.alli + Year))
```


---
\frametitle{Analysis}

Climate Alliance states tend to have a better AQI on average but it is not significant.  

This might be because the Climate Alliance only went into effect 3 years ago in 2017. Note that
climate Alliance states have better improvements on AQI on average.

## County Level Effects on AQI
Using the data found by the USDA’s Economic Research Service, we look for predictors in counties to determine air quality and find correlations.
This begins by merging the 2019 AQI with the latest USDA ERS data. We use 2019 data to avoid skewing due to the 2020 West Coast fires.

---

\frametitle{Merging AQI Data with County Data}

To begin the analysis, we start by merging county data with AQI data. We start by
merging all three sets of ERS county data, and then we merge by county and state.  

We only take the data from year 2019 to keep it consistent. We are avoiding using
2020 data due to the fires on the West coast skewing data.

\tiny
``` {r cleanmerge}
county_data_jobs <- fread("data/Release23_June2021/Jobs.csv")
county_data_people <- fread("data/Release23_June2021/People.csv")
county_data_income <- fread("data/Release23_June2021/Income.csv")


county_data_master <- merge(county_data_jobs, county_data_people, by = c("County", "FIPS", "State"))
county_data_master <- merge(county_data_master, county_data_income, by = c("County", "FIPS", "State"))
county_data_master$State <- abbr2state(county_data_master$State)
county_data_master <- county_data_master %>% rename(state = State)

dated_subbed_data <- subbed_data[subbed_data$Year == 2019,] %>% select(state, Year, County, med.aqi, delta.aqi.state)

master_data <- merge(county_data_master, dated_subbed_data, by = c("County", "state"))

glmnet_data <- master_data %>% select(-County, -FIPS) %>% drop_na()

hist(glmnet_data$med.aqi)
```


---
\frametitle{Running the LASSO Algorithm}

Break the cleaned and merged dataset into X and Y for use with cv.glmnet. We use
set.seed(1) for consistency.

\tiny
``` {r glmnet}

set.seed(1)
X <- model.matrix(delta.aqi.state~.-state, glmnet_data)[,-1]
y <- glmnet_data$delta.aqi.state


fit.cv <- cv.glmnet(X, y, alpha = 1, nfolds = 10)

coeffs <- coef(fit.cv)
nonzeros <- rownames(as.matrix(coeffs[which(coeffs != 0), 0]))
non_state_intercept_coef <- nonzeros[2:length(nonzeros)]

select_cols <- c("med.aqi", non_state_intercept_coef)
lm_cols <- glmnet_data %>% select(select_cols)

fit.lm <- lm(data = lm_cols, med.aqi ~ .)

Anova(fit.lm)

```


---
\frametitle{Backwards Selection with Anova}

We remove the variables that are the least relevant to see what factors remain.
\tiny
``` {r anova1, include = FALSE}
fit.lm <- update(fit.lm, .~. -PctEmpMining)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -ForeignBornCaribPct)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -Ed3SomeCollegePct)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpTrans)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -AsianNonHispanicPct2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpManufacturing)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -LandAreaSQMiles2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -UnempRate2012)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PopChangeRate1819)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -UnempRate2020)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -MultipleRacePopChangeRate0010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpAgriculture)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -Immigration_Rate_2000_2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -ForeignBornEuropePct)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpChange1920)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -MultipleRacePct2010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -Net_International_Migration_Rate_2010_2019)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -Deep_Pov_All)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -PctEmpConstruction)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -NetMigrationRate0010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -NaturalChangeRate0010)
Anova(fit.lm)
fit.lm <- update(fit.lm, .~. -NaturalChangeRate0010)
Anova(fit.lm)
```


```{r assign final}
fit.final <- fit.lm
```


---
\frametitle{Examining the Final Fit - Do the Assumptions of the Linear Model Hold Up?}


\tiny
``` {r finalFit, size = "tiny"}

temp <- summary(fit.final)$coefficients[, c("Estimate")]
non_states <- summary(fit.final)$coefficients[50:length(temp)]

non_states

```

From the final model, we see that most of the impact on AQI is geographical. For example, 
the increase from ForeignBornMexNum and NetMigrationNum could signal that states closer to
the Mexican border tend to have worse AQIs due to their location. However, the
most clear predictors are the states themselves.

The assumptions for linearity appear to hold up until about 1 standard deviation below
the mean.


```{r CA Random Forest}
set.seed(467)
ind <- sample(2, nrow(lm_cols), replace = T, prob = c(0.7, 0.3))
train <- lm_cols[ind==1,]
test <- lm_cols[ind==2,]

library(randomForest)
set.seed(200)
rf.fit <- randomForest(med.aqi~., data = train, ntree = 500)

plot(rf.fit, col="red", pch=16, type="p", 
     main="default plot, ")

rf.error.p <- 1:19
for (i in 1:19)
{
  rf.fit <- randomForest(med.aqi~., train, mtry=i, ntree=500)
  rf.error.p[i] <- rf.fit$mse[500]
}

plot(1:19, rf.error.p, pch=16,
     main = "Testing Errors: ntree = 500",
     xlab="mtry",
     ylab="OOB MSE of mtry")
lines(1:19, rf.error.p)

set.seed(250)
rf.fit <- randomForest(med.aqi~., train, mtry=2, ntree=500)
rf.fit.pred <- predict(rf.fit, train)
MSE_test2 <- mean((train$med.aqi - rf.fit.pred)^2)
MSE_test3 <- mean((test$med.aqi - rf.fit.pred)^2)

rf.fit$importance
attributes(rf.fit)

varImpPlot(rf.fit, sort=TRUE, n.var=min(30, nrow(rf.fit$importance)))

plot(rf.fit)
```

---
\frametitle{Diagnostic Plots}

``` {r plots, fig.show="hold", out.width="50%"}
plot(fit.final, 1)
plot(fit.final, 2)
```

## Conclusion

The overall objective of this study was to use the AQI of counties across the USA to determine the impact of variables on the climate. Using data collected by the EPA, we were able to focus on the effect of the Climate Alliance on curbing the deterioration of the AQI across the nation, as well as the correlation between aspects of counties and their air quality.

From this study, we were able to conclude that the Climate Alliance has not had much of an effect yet on the AQI of member states, but do have better AQIs on average compared to other states. We were also able to see that most of the impact on the AQI is geographical based on the significant variables of the model. 

